```{r data generation, echo = FALSE, results = "hide"}

U = sample(c(1.5,3),1)
Q =  sample(1:3,1)
sig.level = sample(c(.05,.10),1)

D = list('major' = 
           c("Math/Science", "Arts and Humanities", "Business and Economics", "Others"),
         'race' = 
           c("White", "Hispanic", "Black", "Other"),
         'party' = 
           c("Democrat", "Republican", "Independent/Other"),
         'residency' = 
           c("Dorm", "Apartment", "Live at Home")
         )

N = lengths(D)

q.dist = 
  lapply(N, function(x) sample(20:100,size = x))
q.dist = 
  lapply(q.dist, function(x) round(x/sum(x),3))
q.dist = 
  lapply(q.dist, function(x) x/sum(x))

sim.data =
  lapply(q.dist, function(x) round(x * sample(20:300,1) ,0))

sim.data =
  lapply(sim.data, function(x) 
    abs(round(x + sample(round(-U*sd(x)):round(U*sd(x)),length(x)) ,0)))

q.data <- data.frame(
  'Catagory' = D[[Q]],
  'Frequency' = sim.data[[Q]]
)
q.distrobution <- data.frame(
  'Catagory' = D[[Q]],
  'Percent' = round(q.dist[[Q]]*100,2)
)
observed = q.data$Frequency
expected = q.dist[[Q]]

test.out <- chisq.test(x = observed, p = expected)
test.reject <- test.out$p.value <= sig.level

if(test.reject){
  q.ans <- c(0,0,1,0)
  q.solution <- q.ans == 1
} else{
  q.ans <- c(0,1,0,0)
  q.solution <- q.ans == 1
}  
  
q.text <- 
      list('major' = 
           "The distribution of majors at the college is shown in Table A. Suppose a random sample of athletes from the college are surveyed. The Table B shows the results of the survey.",
         'race' = 
           "The racial distribution for the state is shown in Table A. You are interested in seeing if the distribution at your college fits the distribution of the state. The Table B shows the results of the survey of students from your college.",
         'party' = 
           "The distribution of political affiliation in California is shown in Table A. Suppose a random sample of Californians who own their own homes are surveyed. The Table B shows the results of the survey.",
         'residency' = 
           "The statewide distribution of type of residency for students is given in Table A. Suppose a random sample of students from your college are surveyed. The Table B shows the results of the survey."
         )




```


Question
========

`r q.text[[Q]]`

**Table A**

`r kable(q.data)`

**Table B**
`r kable(q.distrobution)`

Use the following output to evaluate weather the sample (Table B) fits the population distribution (Table A) at the `r sig.level` level of significance.

```{r, echo=FALSE}
test.out
```

**Hints:**

[Link to Text Book](https://stats.libretexts.org/Bookshelves/Applied_Statistics/Book%3A_Learning_Statistics_with_R_-_A_tutorial_for_Psychology_Students_and_other_Beginners_(Navarro)/12%3A_Categorical_Data_Analysis/12.01%3A_The_2_Goodness-of-fit_Test?contentOnly "LSR - 12.1: The χ2 Goodness-of-fit Test"){target="blank"}

Answerlist
----------
* Reject the null hypothesis, there is not sufficient evidence to conclude that the sample (Table B) does not fit the population distribution (Table A)
* Fail to reject the null hypothesis, there is not sufficient evidence to conclude that the sample (Table B) does not fit the population distribution (Table A)
* Reject the null hypothesis, there is sufficient evidence to conclude that the sample (Table B) does not fit the population distribution (Table A)
* Fail to reject the null hypothesis, there is sufficient evidence to conclude that the sample (Table B) does not fit the population distribution (Table A)

Solution
========

```{r solutionlist, echo = FALSE, results = "asis"}
answerlist(ifelse(q.solution, "True", "False"), markup = "markdown")
```

Meta-information
================
extype: schoice
exsolution: `r mchoice2string(q.solution, single = TRUE)`
exname: chi2goodfit


