```{r data generation, echo = FALSE, results = "hide"}

U = sample(c(TRUE,FALSE),1)
N = c(4,6)
Q =  sample(1:2,1)
sig.level = sample(c(.05,.10),1)

sim.data <- function(n, uniform = T){
   if(uniform){ 
    sample(10:100,1) %>% 
    rep(n) +
    sample(-10:20,n,replace = F) %>% 
    return()
   } else {
    sample(30:70,n) %>% 
    return()  
   }
}

D = list('cards' = c("Hearts", "Diamonds", "Spades", "Clubs"),
         'dice' = paste(1:6))
q.data <- data.frame(
  'Day' = D[[Q]],
  'Frequency' = sim.data(n = N[Q],uniform = U)
)
observed = q.data$Frequency

q.text <- list(
'cards' = 
  paste0("You want to see if a card dealer is favoring one suit over another. You observe the dealer pick a card, put it back in the deck, shuffle, and then repeat the process ", 
fmt(sum(observed), 0),
" times. The results are displayed in the table below."),

'dise' = 
  paste0("A die is rolled ",
fmt(sum(observed), 0),
" times to see if it is fair. The table below shows the outcomes.")
)

test.out <- chisq.test(observed)
test.reject <- test.out$p.value <= sig.level

if(test.reject){
  q.ans <- c(0,0,1,0)
  q.solution <- q.ans == 1
} else{
  q.ans <- c(0,1,0,0)
  q.solution <- q.ans == 1
}

```


Question
========

`r q.text[[Q]]`

`r kable(q.data)`

Use the following output to evaluate weather the above random process is fair at the `r sig.level` level of significance.

```{r, echo=FALSE}
test.out
```

**Hints:**

[Link to Text Book](https://stats.libretexts.org/Bookshelves/Applied_Statistics/Book%3A_Learning_Statistics_with_R_-_A_tutorial_for_Psychology_Students_and_other_Beginners_(Navarro)/12%3A_Categorical_Data_Analysis/12.01%3A_The_2_Goodness-of-fit_Test?contentOnly "LSR - 12.1: The χ2 Goodness-of-fit Test"){target="blank"}

Answerlist
----------
* Reject the null hypothesis, there is not sufficient evidence to conclude that the process is unfair
* Fail to reject the null hypothesis, there is not sufficient evidence to conclude that the process is unfair
* Reject the null hypothesis, there is sufficient evidence to conclude that the process is unfair
* Fail to reject the null hypothesis, there is sufficient evidence to conclude that the process is unfair





Solution
========

```{r solutionlist, echo = FALSE, results = "asis"}
answerlist(ifelse(q.solution, "True", "False"), markup = "markdown")
```




Meta-information
================
extype: schoice
exsolution: `r mchoice2string(q.solution, single = TRUE)`
exname: chi2uniform2


