---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
output: 
  html_document:
    output: html_notebook
    css: style.css
    code_folding: show
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 1. How ANOVA Works

This practice follows the example in LSR: 14.3

## **Step 1:** Load the data and calculate descriptive statistics.

```{r}
# Load the data
load(
  url("https://ljkelly3141.github.io/ABE/data/clinicaltrial.Rdata")
  )
head(clin.trial)
```

```{r}
# Calculate the overall mean of `mood.gain`
Y.bar <- mean(clin.trial$mood.gain)
Y.bar

# Count of number subjects in each group 
drug.count <- xtabs( ~drug, clin.trial ) 
drug.count

# Mean of mood.gain by drug group
mood.gain_mean <- aggregate(mood.gain ~ drug, clin.trial, mean) 
mood.gain_mean
```

## **Step 2:** Calculate the total sum of squared difference from the mean

$$\sum_{i=1}^{N}{(Y_i-\bar{Y})^2}$$

```{f}
SS.tot <- sum((clin.trial$mood.gain - Y.bar)^2)
SS.tot
```

## **Step 3:** Within-group sum of squares

$$SS_{w} = \sum_{k=1}^{G}{\sum_{i=1}^{N_k}{(Y_{ik}-\bar{Y_k})^2}}$$

```{r}
# First, we make a new variable that has the group mean of each observation.
clin.trial$group.mean <- mood.gain_mean[clin.trial$drug,2]
head(clin.trial)
```

```{r}
# Then, we can simply sum the squared difference.
SS.w <- sum((clin.trial$mood.gain - clin.trial$group.mean)^2)
SS.w
```

## **Step 4:** Between-group sum of squares

$$SS_{b} = \sum_{k=1}^{G}{\sum_{i=1}^{N_k}{(\bar{Y_k}-\bar{Y})^2}}$$

```{r}
SS.b <- sum((clin.trial$group.mean - mean(clin.trial$mood.gain))^2)
SS.b
```

## **Step 5:** Degrees of freedom

```{r}
# Within-group
DF.w <- nrow(clin.trial)-nrow(mood.gain_mean)
DF.w

# Between-group
DF.b <- nrow(mood.gain_mean)-1
DF.b
```

## **Step 6:** Mean squares value

```{r}
MS.w <- SS.w/DF.w
MS.w

MS.b <- SS.b/DF.b
MS.b
```

## **Step 7:** F-Stat

```{r}
F.stat <- MS.b/MS.w
F.stat
```

## **Step 8:** p-value

```{r}
p.value <- pf( F.stat, df1 = DF.b, df2 = DF.w, lower.tail = FALSE)
p.value
```


# 2. Use the `aggregate()` and `aov()` functions to calculate the average mood gain.

```{r}
# Mean of mood.gain by drug group
mood.gain_mean <- aggregate(mood.gain ~ drug, data = clin.trial, FUN = length) 
mood.gain_mean

# Conduct ANOVA
mood.gain_aov <- aov(mood.gain ~ drug, data = clin.trial)
summary(mood.gain_aov)
```

# 3. Check assumptions

## 3.A Check the homogeneity of variance assumption and run the Welch One-way Test.

```{r}
# Load the car package
library("car")

# Draw a boxplot of one.dose by state.control
boxplot(mood.gain ~ drug, data = clin.trial)

# Conduct the leveneTest()
leveneTest(mood.gain_aov, center=mean)

# Conduct the oneway.test()
mood.gain_welch <- oneway.test(mood.gain ~ drug, data = clin.trial)
mood.gain_welch
```

## 3.B Check the normality of residuals assumption




```{r}
# Plot a histogram of the residuals using hist()
hist(mood.gain_aov$residuals)

# Plot the Normal Q-Q Plot using qqnorm()
qqnorm(mood.gain_aov$residuals)

# Conduct the Shapiro Test using shapiro.test()
shapiro.test(mood.gain_aov$residuals)

# Conduct the Kruskal-Wallis test using kruskal.test()
kruskal.test(mood.gain ~ drug, data = clin.trial)

```

