---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
output: 
  html_document:
    output: html_notebook
    css: style.css
    code_folding: show
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 1. The $\chi^2$ Goodness-of-fit Test

A dice is rolled as follows:

|  1  |  2  |  3  |  4  |  5  |  6  |
|:---:|:---:|:---:|:---:|:---:|:---:|
| 55  | 31  | 37  | 47  | 48  | 32  |

Use the $\chi^2$ goodness of fit test to determine if the dice is fair to a 95% confidence level.

## 1.A Calculate $\chi^2$ goodness of fit test "by hand"

-   **Step 1:** Enter observed frequencies into a vector named `observed`

    ```{r}
    observed <- c(
      "1" = 55,
      "2" = 31,
      "3" = 37,
      "4" = 47,
      "5" = 48,
      "6" = 32
    )
    ```

-   **Step 2:** Save the number of trials in a variable named `N`

    ```{r}
      N <- sum(observed)
    ```

-   **Step 3:** Save the degrees of freedom in a variable named `degrees.freedom`

    ```{r}
    degrees.freedom <- length(observed) - 1
    ```

-   **Step 4:** Save the probabilities of each outcome in a variable named `probabilities`

    ```{r}
    probabilities <- rep(1 / length(observed), length(observed))
    ```

-   **Step 5:** Save the expected value of each outcome in a variable named `expected`

    ```{r}
    expected <- N * probabilities
    ```

-   **Step 6:** Save the test statistic in a variable named `chisq.stat`

    ```{r}
    chisq.stat <- sum((observed - expected) ^ 2 / expected)
    ```

-   **Step 7:** Save the p-value in a variable named `chisq.pvalue`

    ```{r}
    chisq.pvalue <- pchisq(q = chisq.stat, 
                           df = degrees.freedom, 
                           lower.tail = FALSE)
    ```

-   **Step 8:** Save both `chisq.stat` and `chisq.pvalue` in a vector named `chisqr`. Print `chisqr`

    ```{r}
    chisqr <- c("chisq stat" = chisq.stat,
                "chisq pvalue" = chisq.pvalue)
    chisqr
    ```

## 1.B Use the chisq.test() function to repeat the chi squared test.

```{r}
built.in.chisq <- chisq.test(x = observed, p = probabilities)
built.in.chisq
```

## 1.C What about two dice?

### We need the expected probabilities of roiling 1 to 12:

| **ROLL A...** | **PROBABILITY** |
|:--------------|:----------------|
| 2             | 1/36 (2.778%)   |
| 3             | 2/36 (5.556%)   |
| 4             | 3/36 (8.333%)   |
| 5             | 4/36 (11.111%)  |
| 6             | 5/36 (13.889%)  |
| 7             | 6/36 (16.667%)  |
| 8             | 5/36 (13.889%)  |
| 9             | 4/36 (11.111%)  |
| 10            | 3/36 (8.333%)   |
| 11            | 2/36 (5.556%)   |
| 12            | 1/36 (2.778%)   |

```{r}
probabilities <- c(1,2,3,4,5,6,5,4,3,2,1)
probabilities <- probabilities/sum(probabilities)
probabilities
```

### Now lets roll some dice:

```{r}
n.dice <- 2
n.sides <- 6
rep <- 100

my.dice.sum <- function(n.dice, n.sides){
  dice <- sample(1:n.sides, size = n.dice, replace = TRUE)
  return(sum(dice))
}
observed <- replicate(rep, my.dice.sum(n.dice, n.sides))

observed <- table(cut(observed, 2:13, right=FALSE))
observed
```

### Now run the test:

```{r}
built.in.chisq <- chisq.test(x = observed, p = probabilities)
built.in.chisq
```

### Now let's see if the distribution is the same for three dice:

```{r}
n.dice <- 3
n.sides <- 6
rep <- 100

observed <- replicate(rep, my.dice.sum(n.dice, n.sides))
observed <- table(cut(observed, 2:13, right=FALSE))

built.in.chisq <- chisq.test(x = observed, p = probabilities)
built.in.chisq
```

# 2. The $\chi^2$ Test of Independence (or Association)

Are hair color and body type independent? The table below shows the results of a researcher's observations of randomly selected people.

|                   | Blonde | Brunette | Red |
|:------------------|:-------|:---------|:----|
| Short and Slender | 53     | 127      | 31  |
| Short and Pudgy   | 62     | 134      | 36  |
| Tall and Slender  | 71     | 118      | 40  |
| Tall and Heavy    | 46     | 100      | 42  |

Use the $\chi^2$ test for independence to determine if there is a significant (95% confidence level) relationship between body type and hair color. Use the following steps:

-   **Step 1.** Load the data. We need to use the `matrix()` function to input the data. Name the matrix `HairByType`. Here is an example of how to do this:

    ```{r}

    HairByType <- matrix(c(53,127,31,62,134,36,71,118,40,46,100,42),
                         ncol = 3,
                         byrow = TRUE) 

    rownames(HairByType) <- c("Short and Slender","Short and Pudgy","Tall and Slender","Tall and Heavy")

    colnames(HairByType) <- c("Blonde","Brunette","Red")

    HairByType
    ```

-   **Step 2.** Run the test. Store the results in `Test.Stat`. Example code: `Test.Stat <- chisq.test(HairByType)`

    ```{r}
    Test.Stat <- chisq.test(HairByType)
    Test.Stat
    ```

## Let't try a more realistic example

-   Convert the `HairByType` to raw data

    ```{r}
    library(tidyverse)
    Hair.BodyType <- as.data.frame(HairByType) %>% 
      rownames_to_column("Body Type") %>% 
      pivot_longer(-`Body Type`,names_to = "Hair Type") %>% 
      uncount(value)
    head(Hair.BodyType)
    ```

-   Use the raw data to conduct the $\chi^2$ test for independence. To use the `chisqr.test()` function, the data need to be processed using the `table()` function.

    ```{r}
    HairBody.Table <- table(Hair.BodyType)
    HairBody.Table
    ```

-   Now, conduct the test.

    ```{r}
    Test.Stat <- chisq.test(HairBody.Table)
    Test.Stat
    ```

# 3. Corrections and Assumptions

## 3.A The Continuity Correction

-   The Yeats Correction

    ```{r}
    Test.Stat <- chisq.test(HairBody.Table, correct = TRUE)
    Test.Stat

    Test.Stat <- chisq.test(HairBody.Table, correct = FALSE)
    Test.Stat
    ```

-   An example with a 2 x 2 contingency table: Subset of Hair

    ```{r}
    Hair.Table <- Hair.BodyType %>% 
      filter(`Hair Type` != "Red" & str_detect(`Body Type`, "Tall")) %>% 
      table()
    Hair.Table
    
    Test.Stat <- chisq.test(Hair.Table, correct = TRUE)
    Test.Stat

    Test.Stat <- chisq.test(Hair.Table, correct = FALSE)
    Test.Stat
    ```

## 3.B Effect Size

- Calculate Cramer's V

  ```{r}
  HairBody.chisq <- chisq.test(HairBody.Table)

  chisq <- HairBody.chisq$statistic
  N <- sum(HairBody.chisq$observed)
  k <- min(nrow(HairBody.chisq$observed),ncol(HairBody.chisq$observed))
  V <- as.vector(sqrt(chisq/(N*(k-1))))
  V
  ```

- Just for fun...Let's make a function

  ```{r}
  cramersV <- function(cs.test){
    chisq <- cs.test$statistic
    N <- sum(cs.test$observed)
    k <- min(nrow(cs.test$observed),ncol(cs.test$observed))
    return(as.vector(sqrt(chisq/(N*(k-1)))))
  }
  
  cramersV(HairBody.chisq)
  ```


## 3.C Assumptions of the Test(s)

- Expected frequencies are sufficiently large. 
- Data are independent of one another.

# 4. The $\chi^2$ Test in Practice

## 4.A Goodness of Fit Test from Raw Data

This section follows the example presented in LSR: 12.1.1

```{r}
# Load the data
load(
  url("https://ljkelly3141.github.io/ABE/data/randomness.Rdata")
  )
head(cards)

# Tabulate the observed table
observed <- table(cards$choice_1)
observed

# Run the test
Test.Stat <- chisq.test(observed) 
Test.Stat
```

## 4.B Fisher Exact Test 

This section follows the example presented in LSR: 12.7

```{r}
# Load the data
load(
  url("https://ljkelly3141.github.io/ABE/data/salem.Rdata")
  )
head(trial)

# Tabulate the observed table
salem.tabs <- table(trial)
salem.tabs

# Run the test chi squared test. Note that you are warned that the sample is too small.
chisq.stat <- chisq.test(salem.tabs) 
chisq.stat

# Run Fisher Exact Test 
fisher.stat <- fisher.test(salem.tabs)
fisher.stat
```

## 4.C The McNemar Test

This section follows the example presented in LSR: 12.8

```{r}
# Load data
load(
  url("https://ljkelly3141.github.io/ABE/data/agpp.Rdata")
)
head(agpp)
summary(agpp) 

# Tabulate the contingency table
right.table <- xtabs( ~ response_before + response_after, data = agpp)
right.table

# Conduct the test
Test.Stat <- mcnemar.test(right.table)
Test.Stat
```

