---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
output: 
  html_document:
    output: html_notebook
    css: style.css
    code_folding: show
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 1. ggplot Basics

## 1.A Load and view the data

```{r}
pacman::p_load("tidyverse")
data(BOD)
head(BOD)
```

## 1.B Set data and mapping for the plot

```{r}
ggplot(data=BOD,
       mapping=aes(x=Time,
                   y=demand))
```

### Pipeing

```{r}
# Load the data
load(
  url("https://ljkelly3141.github.io/ABE/data/parenthood.Rdata")
  )
head(parenthood)

summary(parenthood)

regression.1 <- lm(formula = dan.grump ~ dan.sleep,  data = parenthood)
summary(regression.1)


"https://ljkelly3141.github.io/ABE/data/parenthood.Rdata" %>% 
  url() %>% 
  load() 

```


## 1.C Add a Point Geometry

```{r}
BOD %>% 
ggplot(aes(x=Time, y=demand)) +
  geom_point(size = 5) 
```

## 1.D Add a Line Geometry and some Style

```{r}
ggplot(data=BOD,
       mapping=aes(x=Time,
                   y=demand)) +
  geom_point(size = 5) +
  geom_line(color="red")
```

# 2. Anouther Example

## 2.A Load and view the data

```{r}
pacman::p_load("tidyverse")
data(CO2)
head(CO2)
```

## 2.B Set up a Basic Scaterplot 
```{r}
CO2 %>% ggplot(aes(x = conc,
                   y = uptake,
                   color = Treatment))+
  geom_point()
```

## 2.C Set add Smothing Geometry and some Style

```{r}
CO2 %>% ggplot(aes(x = conc,
                   y = uptake,
                   color = Treatment))+
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth()
```

## 2.D Specify Linear Method of Smoothing and add a Facet

```{r}
CO2 %>% ggplot(aes(x = conc,
                   y = uptake,
                   color = Treatment))+
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(method = lm,se = F) +
  facet_wrap(~Type)
```

## 2.E Add in labels and a theme

```{r}
CO2 %>% ggplot(aes(x = conc,
                   y = uptake,
                   color = Treatment))+
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(method = lm, se = F) +
  facet_wrap(~Type) +
  labs(title = "Consintration of CO2") +
  theme_bw()
```

# 3. Boxplots

## 3.A Set up a Basic Boxplot 

```{r}
CO2 %>% ggplot(aes(x = Treatment,
                   y = uptake )) +
  geom_boxplot()
```

## 3.B Set up a Basic Boxplot 
```{r}
CO2 %>% ggplot(aes(x = Treatment,
                   y = uptake )) +
  geom_boxplot() +
  geom_point()
```

## 3.C Set up a add Point Geometry

```{r}
CO2 %>% ggplot(aes(x = Treatment,
                   y = uptake )) +
  geom_boxplot() +
  geom_point(aes(size = conc,
                 color = Plant))
```

## 3.D Set up a add Point Geometry

```{r}
CO2 %>% ggplot(aes(x = Treatment,
                   y = uptake )) +
  geom_boxplot() +
  geom_point(
    alpha = 0.5,
    aes(size = conc,
        color = Plant))
```

## 3.E Flip the Plot on its Side

```{r}
CO2 %>% ggplot(aes(x = Treatment,
                   y = uptake )) +
  geom_boxplot() +
  geom_point(
    alpha = 0.5,
    aes(size = conc,
        color = Plant)) +
  coord_flip()+
  theme_bw()
```

## 3.F Add a Facet to the Plot

```{r}
CO2 %>% ggplot(aes(x = Treatment,
                   y = uptake )) +
  geom_boxplot() +
  geom_point(
    alpha = 0.5,
    aes(size = conc,
        color = Plant)) +
  coord_flip() +
  theme_bw() +
  facet_wrap(~Type) +
  labs(title = "Chilled vs. Non-chilled")
```

# 4. Scaterplot

## 4.A Set up a Basic Scaterplot 

```{r}
data(mpg)
head(mpg)

mpg %>% ggplot(aes(x = displ,
                   y = cty )) +
  geom_point()
```

## 4.B Add Factors to the Scaterplot 

```{r}
mpg %>% ggplot(aes(x = displ,
                   y = cty )) +
  geom_point(aes(color = drv,
                 size = trans),
             alpha = 0.5)
```

## 4.C Add Smoothing and a Facet

```{r}
mpg %>% ggplot(aes(x = displ,
                   y = cty )) +
  geom_point(aes(color = drv,
                 size = trans),
             alpha = 0.5) +
  geom_smooth(method = lm) + 
  facet_wrap(~year, nrow=1)
```

## 4.D Filter out Outliers

```{r}
mpg %>% 
  filter(cty < 25) %>% 
  ggplot(aes(x = displ,
                   y = cty )) +
  geom_point(aes(color = drv,
                 size = trans),
             alpha = 0.5) +
  geom_smooth(method = lm) + 
  facet_wrap(~year, nrow=1) 
```

## 4.E Add Lables

```{r}
mpg %>% 
  filter(cty < 25) %>% 
  ggplot(aes(x = displ,
                   y = cty )) +
  geom_point(aes(color = drv,
                 size = trans),
             alpha = 0.5) +
  geom_smooth(method = lm) + 
  facet_wrap(~year, nrow=1) +
  labs(x = "Engine Size",
       y = "City MPG",
       title = "Fuel Efficiency")+
  theme_bw()
```

# 5. Visualizing Categorical Data

## 5.A Load and view the data

```{r}
pacman::p_load("tidyverse")
data("msleep")
head(msleep)
```

## 5.B Single Categorical Variable

```{r}
msleep %>% 
  drop_na(vore) %>% 
  ggplot(aes(x = vore)) +
  geom_bar(fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Vore",
       y = NULL,
       title = "Number of Ovservations per Order")
```

## 5.C Comment out `drop_na(vore) %>% `

```{r}
msleep %>% 
# drop_na(vore) %>% 
  ggplot(aes(x = vore)) +
  geom_bar(fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Vore",
       y = NULL,
       title = "Number of Ovservations per Order")
```


## 5.D Dropping too much data

```{r}
msleep %>% 
  drop_na() %>% 
  ggplot(aes(x = vore)) +
  geom_bar(fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Vore",
       y = NULL,
       title = "Number of Ovservations per Order")
```

## 5.E Flip the coordinates 

```{r}
msleep %>% 
  drop_na(vore) %>% 
  ggplot(aes(x = vore)) +
  geom_bar(fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Vore",
       y = NULL,
       title = "Number of Ovservations per Order") +
  coord_flip()
```

## 5.F Reorder by Number of Observations 

```{r}
msleep %>% 
  drop_na(vore) %>% 
  ggplot(aes(x = fct_infreq(vore))) +
  geom_bar(fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Vore",
       y = NULL,
       title = "Number of Ovservations per Order")
```

# 6.  Single Numeric Variable

## 6.A Basic Histogram Setup

```{r}
msleep %>% 
  ggplot(aes(x = awake)) +
  geom_histogram(binwidth = 2, fill = "#97B3C6") +
  theme_bw() +
  labs(x = "Total Sleep",
       y = NULL,
       title = "Histogram of Total Sleep")
```

## 6.B Omit the Histogram Options 

```{r}
msleep %>% 
  ggplot(aes(x = awake)) +
  geom_histogram() +
  theme_bw() +
  labs(x = "Total Sleep",
       y = NULL,
       title = "Histogram of Total Sleep")
```

## 6.C Histograms with Various Bin Widths 

::: {.info-box .note}
I cheated a bit here. The package gridExtra let's me layout multiple plots together. This also demonstrates the power of the grammar of graphics. I can save a basic plot object and add what changes. In this case, the bin width and the title.
:::

```{r, fig.width=8, fig.height=5}
pacman::p_load("gridExtra")

h <- msleep %>% 
  ggplot(aes(x = awake)) +
  theme_bw() +
  labs(x = "Total Sleep",
       y = NULL)

grid.arrange(h + geom_histogram(binwidth = 1, fill = "#97B3C6") + ggtitle("Bin = 1"),
             h + geom_histogram(binwidth = 2, fill = "#97B3C6") + ggtitle("Bin = 2"),
             h + geom_histogram(binwidth = 5, fill = "#97B3C6") + ggtitle("Bin = 5"),
             h + geom_histogram(binwidth = 10, fill = "#97B3C6") + ggtitle("Bin = 10"), 
             nrow = 2)
```

# 7. Visualizing Mixed Data Sets

## 7.A Load the `starwars` data set

```{r}
data(starwars)
head(starwars)
```

## 7.B Barplot with two catagorical variables 

```{r}
starwars %>% 
  filter(hair_color == "black" |
         hair_color == "brown") %>% 
  drop_na(sex) %>% 
  ggplot(aes(hair_color, fill = sex)) +
  geom_bar(position = "dodge",
           alpha = 0.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Sex and Hair Color",
       x = "Hair Color",
       y = "Number")
```

## 7.C Barplot with `position = "fill"`

```{r}
starwars %>% 
  filter(hair_color == "black" |
         hair_color == "brown") %>% 
  drop_na(sex) %>% 
  ggplot(aes(hair_color, fill = sex)) +
  geom_bar(position = "fill",
           alpha = 0.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Sex and Hair Color",
       x = "Hair Color",
       y = "Number")
```

## 7.D Barplot with no Theme

```{r}
starwars %>% 
  filter(hair_color == "black" |
         hair_color == "brown") %>% 
  drop_na(sex) %>% 
  ggplot(aes(hair_color, fill = sex)) +
  geom_bar(position = "dodge",
           alpha = 0.5)
```

## 7.E Barplot with no position specified

```{r}
starwars %>% 
  filter(hair_color == "black" |
         hair_color == "brown") %>% 
  drop_na(sex) %>% 
  ggplot(aes(hair_color, fill = sex)) +
  geom_bar(alpha = 0.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Sex and Hair Color",
       x = "Hair Color",
       y = "Number")
```





