---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
output: 
  html_document:
    code_download: true
    css: style.css
---

# Case Study: Wine Quality

- **Dataset**: [Wine Data](https://ljkelly3141.github.io/ABE-Book/data/winequality-red.csv)
- **Source**: Cortez, Paulo, Cerdeira, A., Almeida, F., Matos, T., and Reis, J.. (2009).[^50-linear_wine-1]
  
[^50-linear_wine-1]: Cortez et al. (2009). UCI Machine Learning Repository. https://doi.org/10.24432/C56S3T.

## Case Background
- Goal: Understand factors contributing to wine quality.
- Dataset: Contains physicochemical attributes and quality ratings.
  
## Description of the Data
- Data is about red wine samples from North Portugal.
- Variables include:

| Variable             | Description                                               |
|----------------------|-----------------------------------------------------------|
| fixed acidity        | Most acids involved with wine or fixed or nonvolatile. |
| volatile acidity     | Amount of acetic acid in wine. |
| citric acid          | Adds freshness and flavor to wines. |
| residual sugar       | Amount of sugar post fermentation. |
| chlorides            | Amount of salt in the wine. |
| free sulfur dioxide  | Prevents microbial growth and oxidation of wine. |
| total sulfur dioxide | Overall SO2, felt in nose and taste. |
| density              | Correlates with alcohol percentage and sugar content. |
| pH                   | Measures wine's acidity on a scale from 0-14. |
| sulphates            | Wine additive affecting SO2 levels. |
| alcohol              | Alcohol content of the wine. |
| quality              | Score between 0 and 10 by experts. | 

## Step 1: Load and Explore Data
- **Loading Libraries**

   - `psych`: For psychometric theory and data analysis.
   - `jtools`: Functions for regression and plotting.
   - `tidyverse`: Collection of R packages for data manipulation and visualization.

  
  ```{r}
  pacman::p_load(psych, jtools, tidyverse)
  ```

- **Loading the Dataset**

  ```{r}
  wine_data <- 
    read.csv("https://ljkelly3141.github.io/ABE-Book/data/winequality-red.csv", 
      header=TRUE, 
      sep=";")
  ```

- **Standardizing the Dataset**

  ```{r}
  wine_data[,names(wine_data) != "quality"] <- standardize(wine_data[,names(wine_data) != "quality"])
  ```

- **Visualizing Data**
  
  ```{r, fig.height=10}
  wine_data %>%
    pivot_longer(cols = -quality, names_to = "Variable", values_to = "Attributes") %>%
    ggplot(aes(y=quality, x=Attributes)) +
    geom_point() +
    geom_smooth(se=FALSE) +
    facet_wrap(~Variable, ncol = 3)
  ```

## Step 2: Linear Regression Model

  ```{r}
  wine.quality_all <- lm(quality ~ ., data = wine_data)
  summ(wine.quality_all)
  ```

## Step 3: Model Refinement
- Iteratively remove irrelevant variables using the `step` function.
  
  ```{r}
  wine.quality_final <- step(wine.quality_all, trace=0)
  summ(wine.quality_final)
  ```

## Step 4: Assessing Model for Nonlinearity
- **Loading Packages**

  ```{r}
  pacman::p_load(tidyverse)
  ```

- **Creating a Data Frame for Plotting** 
  
  ```{r}
  res.fit_data <- data.frame(check.names = FALSE, 
                             'Residuals' = wine.quality_final$residuals, 
                             'Fitted Values' = wine.quality_final$fitted.values)
  ```

- **Creating the Scatter Plot**

  ```{r}
  res.fit_data %>% 
    ggplot(aes(y=`Residuals`, x=`Fitted Values`)) + 
    geom_point() + geom_smooth(se=FALSE) + 
    geom_abline(intercept = 0, slope = 0)
  ```

## Step 5: Test Normality of Residuals

  ```{r}
  hist(wine.quality_final$residuals)
  ```

## Step 6: Test for Heteroscedasticity
- Using the `bptest` function.

  ```{r}
  pacman::p_load(lmtest)
  bptest(wine.quality_final)
  summ(wine.quality_final, digits=4, robust=TRUE)
  ```

## Step 7: Interpretation of the Model
- Linear regression model derived from dataset with 1,599 observations.
- Model utilized the Ordinary Least Squares (OLS) method.
- Variable coefficients and their implications:
  - Intercept: 5.6360
  - `volatile.acidity`: -0.1813
  - `chlorides`: -0.0950
  - `alcohol`: 0.3083
  
