---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
output: 
  html_document:
    code_download: true
    css: style.css
---

# Case Study: Chi-Squared Goodness of Fit Test

## Introduction

This case study explores the application of the chi-squared goodness of fit test on data sourced from the direct marketing campaigns of a Portuguese banking institution.

**Data Source**: [Bank Market Data](https://ljkelly3141.github.io/ABE-Book/data/bank-full.csv) provided by S. Moro, P. Cortez, and P. Rita (2014)[^20-chi_square-1].

### Case Background

- Objective: Verify if the bank's client job distribution mirrors the general job distribution in society.
- Importance: Reflects the diversity and appeal of the bank's services.

### Data Description

Variables of focus from the dataset:

| Variable  | Description                      |
|-----------|----------------------------------|
| age       | Age of the client                |
| job       | Type of job                      |
| marital   | Marital status                   |
| education | Level of education               |
| default   | Has credit in default?           |
| balance   | Average yearly balance, in euros |
| ...       | ...                              |

- The 'job' variable is categorical with 12 categories.

## **Step 1: Load the Data and Inspect the 'job' Variable**

```{r}
# Load Packages
pacman::p_load("tidyverse")

# Load the dataset
bank <- read.csv("https://ljkelly3141.github.io/ABE-Book/data/bank-full.csv", sep = ';')

# Examine 'job' distribution
job.table <- table(bank$job)
as.data.frame(job.table) %>% 
  rename("Job Categories" = "Var1") 
```

## **Step 2: Hypotheses Formulation**

- Null hypothesis ($H_0$): Observed job distribution = Expected job distribution
- Alternate hypothesis ($H_a$): Observed job distribution ≠ Expected job distribution

Mathematically:

- $H_0$: All proportions $p_i = 1/k$
- $H_a$: At least one proportion $p_i \not= 1/k$

## **Step 3: Expected Frequencies and Degrees of Freedom**

1. Determine number of job categories:
   
```{r}
n.categories <- length(job.table)
n.categories
```

2. Compute expected count for each job category:

```{r}
expected.counts <- c(4521, 11303, 1356, 1356, 9042, 2261, 1356, 4521, 904, 6782, 1356, 453)
names(expected.counts) <- names(job.table)
```

## **Step 4: Calculate the Chi-Square Statistic**

```{r}
observed.counts <- table(bank$job)
chi.sq <- sum((observed.counts - expected.counts)^2 / expected.counts)
chi.sq
```

## **Step 5: Determine the P-value**

```{r}
p.value <- pchisq(chi.sq, df = n.categories - 1, lower.tail = FALSE)
p.value
```

## **Step 6: Result Interpretation**

Given the p-value, if it's less than a significance level (e.g., 0.05), then we reject $H_0$. This would indicate that the bank's client job distribution does not reflect the expected distribution.

## References

[^20-chi_square-1]: Moro, S., Cortez, P., & Rita, P. (2014). A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31. Retrieved from [https://archive.ics.uci.edu/ml/datasets/bank+marketing](https://archive.ics.uci.edu/ml/datasets/bank+marketing)

