---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
output: 
  html_document:
    code_download: true
    css: style.css
---


# Case Study: Linear Regression Modeling of House Prices in Windsor

In this case study, we will utilize the `Housing` dataset from the `Ecdat` package to model house prices in the City of Windsor using linear regression. The objective is to determine which variables most significantly impact house prices.

## Loading Necessary Libraries and Dataset

- Libraries used:
    + `tidyverse` for data manipulation and visualization.
    + `Ecdat` for accessing the `Housing` dataset.
    + `psych` for generating descriptive statistics.
    + `jtools` for summarizing linear models.

```{r}
pacman::p_load(tidyverse, Ecdat, psych, jtools)
```

## Description of the Data

- The `Housing` dataset contains data about various properties in the City of Windsor, Canada.
- Each row corresponds to a house.
- Includes variables such as lot size, number of bedrooms, and price of the house.

```{r}
data(Housing)
head(Housing)
```

## Step 1: Explore the Data

### Summary Statistics

- Use the `describe` function from the `psych` package for an overview.

```{r}
describe(Housing, skew = FALSE,ranges = FALSE,omit = TRUE)
```

### Visualizing Relationships

- Scatter plots to visualize relationships between numeric variables and housing price.

```{r}
Housing %>%
  pivot_longer(cols = -c(price,!where(is.numeric)), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=value, y=price, color =fullbase)) +
    geom_point() +
    facet_wrap(~variable, scales="free") +
    geom_smooth(method = "lm",se = FALSE) +
    theme_minimal() +
    labs(title = "Scatter plots of Numeric Variables vs. Housing Price")
```

## Step 2: Estimate Initial Model

- Use the `lm` function to estimate a model using all variables to predict the house price.

```{r}
model_initial <- lm(price ~ .*fullbase, data = Housing)
summ(model_initial)
```

## Step 3: Variable Selection

- Use the `step` function to remove all non-significant variables at the 10% significance level.

```{r}
model_stepwise <- step(model_initial, trace = 0)
summ(model_stepwise)
```

## Step 4: Plot Actual vs. Fitted

- Visualize how well the model’s predictions match the actual prices.

```{r}
predicted_prices <- predict(model_stepwise)
ggplot(Housing, aes(x = price, y = predicted_prices)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE, color = "black") +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_minimal() +
  labs(title = "Actual vs. Fitted Prices", x = "Actual Price", y = "Fitted Price")
```

## Step 5: Re-estimate with a Log-Log Model

- A log-linear model is useful when the underlying relationship appears to be nonlinear.
- Take the natural log of both `price` and `lotsize`.

```{r}
model_initial_log <- lm(log(price) ~ (. - lotsize + log(lotsize)) * fullbase, data = Housing)
model_log <- step(model_initial_log, trace = 0)
summ(model_log)
```

```{r}
predicted_prices <- predict(model_log)
ggplot(Housing, aes(x = log(price), y = predicted_prices)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE, color = "black") +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_minimal() +
  labs(title = "Actual vs. Fitted Prices", x = "Actual Price", y = "Fitted Price")
```

## Step 6: Interpret the Model

- Interpret coefficients in terms of percentage changes.
- Examples of interpretation:
    + For every additional **bedroom**, the house price is expected to increase by approximately 3%.
    + Each additional **bathroom** is associated with a 20% increase in the house price.
    + Having a **driveway** boosts the house price by about 11%.
    + Houses with a **full basement** see a 19% price hike.
    + **Air conditioning** contributes to a 17% price rise.
    + Being in a **preferred area** is associated with a 14% price premium.
    + A 1% increase in **lot size** leads to a 0.29% increase in the house price.
- Interactions like **bathrms:fullbaseyes** and **fullbaseyes:gashwyes** denote the change in the price percentage due to the combined effect of both features.

### Conclusion

- The model's coefficients give insights into the relative importance and effect of different house features on its price.
- Features like the number of bathrooms, air conditioning, and location in a preferred area have substantial impacts.
```
