---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
output: 
  html_document:
    code_download: true
    css: style.css
---

# Case Study: Chi-Squared Test of Independence with Banking Data

## Introduction

This study examines the relationship between a client's job type and possession of a housing loan using a chi-squared test of independence.

**Data Source**: [Bank Market Data](https://ljkelly3141.github.io/ABE-Book/data/bank-full.csv) provided by S. Moro, P. Cortez, and P. Rita (2014)[^20-chi_square-1].

**Objective**: Investigate if the `job` type of a client informs us about their likelihood of having a housing loan.

**Variables of Interest**:
  - `job`: Type of job.
  - `housing`: Possession of housing loan (yes/no).

1. **Load the Data & Visualize with Mosaic Plot**
   - Use the provided link to download and load the data.
   - Examine the distribution of the `job` and `housing` variables.
```{r}
# Loading the dataset
bank <- read.csv("https://ljkelly3141.github.io/ABE-Book/data/bank-full.csv", sep = ";")
# Distribution check
table(bank$job, bank$housing) %>% knitr::kable()
```
   
2. **Formulate a Hypothesis**
   - **Null hypothesis** ($H_0$): `job` and `housing` are independent.
   - **Alternate hypothesis** ($H_a$): `job` and `housing` are dependent.

3. **Conduct the Chi-Squared Test of Independence**
   - Use the `table` function to form a cross tabulation of `job` and `housing`.
```{r}
cross.tab <- table(bank$job, bank$housing)
cross.tab
```
   - Implement the chi-squared test using `chisq.test()` in R.
```{r}
chisq.stat <- chisq.test(cross.tab)
chisq.stat
```

4. **Interpret the Result**
   - If the p-value < $alpha$, reject $H_0$.
   - In this study, due to a very small p-value, $H_0$ is rejected.
   - **Conclusion**: `job` and `housing` variables are dependent.

#### **Final Thoughts**
- The findings suggest certain job categories might be more or less likely to have housing loans.
- Such insights can be valuable for targeted marketing strategies.

## References

[^20-chi_square-1]: Moro, S., Cortez, P., & Rita, P. (2014). A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31. Retrieved from [https://archive.ics.uci.edu/ml/datasets/bank+marketing](https://archive.ics.uci.edu/ml/datasets/bank+marketing)

