---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
output: 
  html_document:
    code_download: true
    css: style.css
---

# Lecture Notes: Chapter 3 - A First Look at R

## Overview
- Introduction to the basic syntax of R
- Simple data analysis using the `mtcars` dataset
- The dataset contains information on different cars from the 1970s.

## 3.1 Understanding Data Frames
- Definition:
  - A two-dimensional data structure similar to a spreadsheet table.
  - Consists of rows (observations) and columns (variables).

- Key points:
  - **Rows**: Represent a single entity or observation.
    - In `mtcars`, each row represents a unique car model.
  - **Columns**: Represent a distinct attribute of an observation.
    - Example: `mpg` column for miles per gallon.
  - Data in columns must be of the same type, but columns can have different types.

- Loading the `mtcars` dataset:
```{r}
data(mtcars)
```

- Loading an Rdata file
```{r}
load(
  url("https://ljkelly3141.github.io/Managerial-Statistics/data/mtcars.Rdata")
  )
```

- Loading a csv file
```{r}
mtcars <- 
  read.csv("https://ljkelly3141.github.io/Managerial-Statistics/data/mtcars.csv")
```

## 3.2 Understanding the `$` Operator in R
- Used to access specific columns from a data frame.
  - Example: 
```{r}
mtcars$mpg
```

## 3.3 Summarizing the miles per gallon variable

- `summary()` function: Provides statistical properties: min, 1st quartile, median, mean, 3rd quartile, max.

```{r}
summary(mtcars$mpg)
```

## 3.3.1 Plotting miles per gallon vs horsepower

- Using the `plot()` function.
  - Example: 
  
```{r}
plot(x = mtcars$hp, 
     y = mtcars$mpg, 
     xlab = "Horsepower", 
     ylab = "Miles per Gallon", 
     main = "Miles per Gallon vs Horsepower")
```

- Parameters:
  - `xlab`: Label for the x-axis.
  - `ylab`: Label for the y-axis.
  - `main`: Main title of the plot.

## 3.4 Fitting a Trend Line
- Introduction to the `lm()` function for linear regression.
  - Example: 
  
```{r}
fit <- lm(mpg ~ hp, data = mtcars)
```

- Visualizing the trend line using the `abline()` function.
  - Example: 
  
```{r}
plot(x = mtcars$hp, 
     y = mtcars$mpg, 
     xlab = "Horsepower", 
     ylab = "Miles per Gallon", 
     main = "Miles per Gallon vs Horsepower")

abline(fit, col="red")
```

## 3.5 Complete Chapter Example

- A recap of all the R code discussed in this chapter.

```{r}
data(mtcars)
mpg_summary <- summary(mtcars$mpg)
mpg_summary
fit <- lm(mpg ~ hp, data = mtcars)
plot(x = mtcars$hp, 
     y = mtcars$mpg, 
     xlab ="Horsepower", 
     ylab = "Miles per Gallon", 
     main = "Miles per Gallon vs Horsepower")
abline(fit, col="red")
```

## Conclusion

- This chapter serves as an introduction to basic operations in R.
- These fundamentals will pave the way for more complex analyses in R.