---
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs/")
  })
institute: "University of Wisconsin-River Falls"
department: "College of Business and Economics"
short-author: "Kelly"
short-date: ''
short-institute: "UWRF"
output: 
  html_document:
    code_download: true
    css: style.css
---

# Case Study: Online Retail II data set

- This study involves the **Online Retail II data set** from the UCI Machine Learning Repository.
- The dataset contains transactions for a UK-based online retail between 01/12/2009 and 09/12/2011.
- **Key Question**: Is there a significant difference in the average transaction amount across different countries?

## Objective

- Conduct an ANOVA to determine the difference in the quantity of products purchased across various countries.
- We aim to manually compute ANOVA without using the built-in `anova()` function.

## Loading Necessary Packages

- Main package for data manipulation: `dplyr`
- It is part of the `tidyverse` collection.

```{r setup, include=FALSE}
pacman::p_load(tidyverse,openxlsx,readxl)
```

## Description of the Data

- The dataset can be loaded and its initial rows visualized using:

```{r load_data}
data <- read.xlsx("https://ljkelly3141.github.io/ABE-Book/data/online_retail_II.xlsx")
head(data)
```

### Variables

- `InvoiceNo`: Unique 6-digit number for each transaction.
- `StockCode`: Unique 5-digit product code.
- `Description`: Product name.
- `Quantity`: Quantity of product per transaction.
- `InvoiceDate`: Date and time of the transaction.
- `UnitPrice`: Product price per unit (£).
- `CustomerID`: Unique 5-digit customer identifier.
- `Country`: Name of the country of the customer.

## Data Cleaning

### Step 1: Remove missing values
- The code removes rows with missing values in 'CustomerID' and 'Description'.

```{r step1}
data_clean <- data %>%
  filter(!is.na(Customer.ID)) %>%
  filter(!is.na(Description))
```

### Step 2: Filter for positive quantities and prices
- Ensure 'Quantity' and 'UnitPrice' are positive.

```{r step2}
data_clean <- data_clean %>%
  filter(Quantity > 0) %>%
  filter(Price > 0)
```

### Step 3: Calculate Total Amount of each Invoice

- Compute the total transaction amount and prepare the data.

```{r}
data_final <- data_clean %>%
  mutate(TransTotal = Price * Quantity) %>%
  group_by(Invoice) %>%
  mutate(InvoiceTotal = sum(TransTotal)) %>%
  distinct(Invoice, .keep_all = TRUE) %>%
  select(Invoice,InvoiceDate,InvoiceTotal,Country)

write.csv(data_final, file = "../docs/data/online_retail_II.csv")
```

## Performing the ANOVA Test Manually

### Step 1: Calculate means
- Get country-wise invoice count and average.

```{r anova_step1}
country.summary <- data_final %>% 
  group_by(Country) %>% 
  summarise(CountryCount = n(),
            CountryMean = mean(InvoiceTotal)) %>% 
  arrange(desc(CountryCount))
```

- Overall mean:

\[\bar{Y} = \frac{1}{N} \sum_{i=1}^{N} Y_i\]

```{r overall_mean}
overall_mean <- mean(data_final$InvoiceTotal)
```

### Step 2: Total Sum of Squares

\[ SS_{\text{tot}} = \sum_{i=1}^{N} (Y_i - \bar{Y})^2 \]

```{r ss_tot}
SS.tot <- sum((data_final$InvoiceTotal-overall_mean)^2)
```

### Step 3: Within-group sum of squares

\[ SS_w = \sum_{k=1}^{G} \sum_{i=1}^{N_k} (Y_{ik} - \bar{Y}_k)^2 \]

```{r ss_w}
data_final <- data_final %>% 
  group_by(Country) %>% 
  mutate(CountryMean = mean(InvoiceTotal)) %>% 
  ungroup()

SS.w <- sum((data_final$InvoiceTotal - data_final$CountryMean)^2)
```

### Step 4: Between-group sum of squares

\[ SS_b = \sum_{k=1}^{G} \sum_{i=1}^{N_k} (\bar{Y}_k - \bar{Y})^2 \]

```{r ss_b}
SS.b <- sum((data_final$CountryMean - overall_mean)^2)
```

### Step 5: Degrees of freedom

\[ DF_w = N - G \]
\[ DF_b = G - 1 \]

```{r df}
DF.w <- nrow(data_final)-nrow(country.summary)
DF.b <- nrow(country.summary)-1
```

### Step 6: Mean squares value

\[ MS_w = \frac{SS_w}{DF_w} \]
\[ MS_b = \frac{SS_b}{DF_b} \]

```{r ms}
MS.w <- SS.w/DF.w
MS.b <- SS.b/DF.b
```

### Step 7: F-Stat

\[ F_{\text{stat}} = \frac{MS_b}{MS_w} \]

```{r f_stat}
F.stat <- MS.b/MS.w
```

### Step 8: p-value

\[ p_{\text{value}} = P(F > F_{\text{stat}}) \]

```{r p_value}
p.value <- pf( F.stat, df1 = DF.b, df2 = DF.w, lower.tail = FALSE)
```

## Conclusion

- The ANOVA results show a significant difference between invoice totals across countries.
